Flexible Probabilistic Forecasting with Distributional Trees and Forests

### 2018-03-31 pkg/disttree/inst/MIP-Seminar/slides.Rnw
revised slides
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2108 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: ee8043bb64140d2bca8e56dc8a484547bc43bd13

\begin{frame}[fragile]
\frametitle{Statistical modeling}
\vspace{0.18cm}
Approximation of reality based on a set of observations.\\
\vspace{0.19cm}
\visible<2, 5-> {Assumption:} 
\vspace{-0.7cm}
\begin{center}
\begin{tikzpicture}
\visible<2, 5->{
\node[] at (0,0.4) {X};
}
\visible<3-4>{
\node[inner sep=0pt] (today) at (-1.8,-0.4)
    {\includegraphics[width=.365\textwidth]{today.png}};
\node[font=\scriptsize] at (-1.9,-2) {15.03.2018};
}
\visible<2->{
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
}
\visible<2, 5->{
\node[] at (3.5,0.4) {Y};
}
\visible<3-4>{
\node[inner sep=0pt] (future) at (5.3,-0.4)
    {\includegraphics[width=.365\textwidth]{future.png}};
\node[font=\scriptsize] at (5.4,-2) {16.03.2018};
}
\end{tikzpicture}
\end{center}
\visible<4>{
\vspace{0.3cm}
X = observed precipitation of today,\\
\qquad current temperature / humidity / \dots,\\
\qquad physical model output\\
\vspace{0.15cm}
Y = daily precipitation sums
}

\vspace{-4cm}
\visible<6->{
\textbf{Application/Motivation:}
\begin{itemize}
\item Prediction.
\item Caption of dependencies and interactions.
\end{itemize}
}
\vspace{0.5cm}
\visible<7->{
\textbf{Two main approaches / ``cultures'':}
\begin{itemize}
\item Data modeling.
\item Algorithmic modeling.
\end{itemize}
}
\end{frame}

\begin{frame}[fragile]
\frametitle{Statistical modeling}
\vspace{0.2cm}
\textbf{Data modeling:}
\vspace{0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {f(X; par)};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\begin{itemize}
\item Approximate data generating process up to a few unknown processes.
\item Mathematical equations, model assumptions.
\item Data: estimation of parameters.
\end{itemize}
\only<1>{\vspace{1.5cm}}

@@

\visible<2->{
\begin{minipage}{0.65\textwidth}
\textbf{Example:} 
\only<2-3>{Linear model}\only<4>{Generalized additive model}\only<5-6>{Generalized additive model\\ for location, scale, and shape}
\\
\only<2-3>{
$$\visible<3>{\mathbb{E}(Y|X) = } f(X; \alpha) = \alpha_0 + \alpha_1 \cdot X$$  
}
\only<4>{
$$\mathbb{E}(Y|X) = f(X; \alpha) = \alpha + s(X)$$  
}
\only<5-6>{
\vspace{-0.8cm}
\begin{align*}
\visible<6>{\mathbb{E}(Y|X) = }\mu(X) = \alpha_{\mu} + s_{\mu}(X)\\  
\visible<6>{\sqrt{Var(Y|X)} = }\sigma(X) = \alpha_{\sigma} + s_{\sigma}(X)
\end{align*}
}
\end{minipage}
%\vspace{-0.6cm}
%\hfill
\begin{minipage}{0.3\textwidth}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<2-3>{
<<plot_motivation_GLM_single, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<4>{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\only<5-6>{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
}
\end{minipage}
}
\end{frame}

@@

sp_rain <- partysplit(2L, index = 1:2)
sp_humidity <- partysplit(4L, breaks = 80)

pn <- partynode(1L, split = sp_rain, kids = list(
  partynode(2L, split = sp_humidity, kids = list(
      partynode(3L, info = "5 mm/24h"),
      partynode(4L, info = "8 mm/24h"))),
  partynode(5L, info = "18 mm/24h")))

p <- party(pn, weather)
@
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
y <- ytrue + rnorm(nobs,0,0.3)
for(i in 1:nobs) ytree[i] <- if(x[i]<1/3) 0.5 else {if(x[i]<2/3) 2 else 1}
@
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@

\begin{frame}[fragile]
\frametitle{Statistical modeling}
\vspace{0.2cm}
\textbf{Algorithmic modeling:}
\vspace{0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\draw[<->] (0.1,0.1) .. controls (0.2, -0.4) and (0.4, -0.5) .. (0.8,-0.5);
\draw[<->] (2.7,-0.5) .. controls (3.1, -0.5) and (3.3, -0.4) .. (3.4,0.1);
\node[font=\scriptsize] at (1.75,-0.5) {algorithm};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\begin{itemize}
\item Flexible approximation with no/few assumptions.
\item Data-driven model fitting.
\end{itemize}
\only<1>{\vspace{1cm}}
\visible<2->{
\begin{minipage}{0.65\textwidth}
%\vspace{-0.5cm}
\textbf{Example:} \only<2>{Tree model}\only<3>{Forest model}\\
%\vspace{-0.8cm}
\begin{center}
\only<2>{
\vspace{-0.7cm}
\resizebox{0.6\textwidth}{!}{
<<plot_raintree, fig=TRUE, echo=FALSE, width=4.2, height=2.6>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.3,0))
paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(p,
     tp_args = list(FUN = identity, width = 9, fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1)
@
}
}
\only<3>{
\resizebox{0.6\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n00) at (1, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n01) at (0.5, 1) {};
  \draw[-, line width=1pt] (n00) -- (n01);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n02) at (1.5, 1) {};
  \draw[-, line width=1pt] (n00) -- (n02);

  \node[ellipse, fill=HighlightBlue!70, align=center] (n10) at (3, 2) {};
  \node[ellipse, fill=HighlightBlue!70, align=center] (n11) at (2.5, 1) {};
  \draw[-, line width=1pt] (n10) -- (n11);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n12) at (3.5, 1) {};
  \draw[-, line width=1pt] (n10) -- (n12);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n13) at (2, 0) {};
  \draw[-, line width=1pt] (n11) -- (n13);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n14) at (2.8, 0) {};
  \draw[-, line width=1pt] (n11) -- (n14);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n15) at (3.2, 0) {};
  \draw[-, line width=1pt] (n12) -- (n15);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n16) at (4, 0) {};
  \draw[-, line width=1pt] (n12) -- (n16);

  \node[ellipse, fill=HighlightBlue!70, align=center] (n20) at (5, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n21) at (4.5, 1) {};
  \draw[-, line width=1pt] (n20) -- (n21);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n22) at (5.5, 1) {};
  \draw[-, line width=1pt] (n20) -- (n22);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n23) at (5, 0) {};
  \draw[-, line width=1pt] (n22) -- (n23);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n24) at (6, 0) {};
  \draw[-, line width=1pt] (n22) -- (n24);
\end{tikzpicture}
}
}
\end{center}
\end{minipage}
\begin{minipage}{0.3\textwidth}
\begin{center}
\only<2>{
\vspace{-0.29cm}
<<plot_motivation_regtree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "grey", lwd=5, main = "")
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
@
}
\only<3>{
<<plot_motivation_randforest, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "grey", lwd=5, main = "")
lines(x = x, y = yforest, col = pal["forest"], lwd=7, main = "")
@
}
\end{center}
\end{minipage}
}
\end{frame}

\begin{frame}
\frametitle{Goals}
\vspace{0.3 cm}
\textbf{Synthesis:} \\
\vspace{0.2 cm}
\begin{itemize}
  \item Probabilistic forecast\\
  $\rightarrow$ Distributional approach
  \medskip
  \item Automatic detection of steps and abrupt changes,\\
  automatic selection of variables and interactions\\
  $\rightarrow$ Tree approach
  \medskip
  \item Smoothing and regularization\\
  $\rightarrow$ Forest approach
\end{itemize}
\end{frame}

>>>>

\begin{frame}[fragile]
\frametitle{Statistical modeling}
\visible<2,3, 6-> {Assumption:}
\vspace{-0.7cm}
\begin{center}
\begin{tikzpicture}
\visible<2,3, 6-> {\node[] at (0,0.4) {X};}
\visible<4-5>{
\node[inner sep=0pt] (today) at (-1.8,-0.4)
    {\includegraphics[width=.365\textwidth]{today.png}};
\node[font=\scriptsize] at (-1.9,-2) {15.03.2018};
}
\visible<2->{
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
}
\visible<2,3, 6->{
\node[] at (3.5,0.4) {Y};
}
\visible<4-5>{
\node[inner sep=0pt] (future) at (5.3,-0.4)
    {\includegraphics[width=.365\textwidth]{future.png}};
\node[font=\scriptsize] at (5.4,-2) {16.03.2018};
}
\end{tikzpicture}
\end{center}
\visible<3, 6->{
\vspace{-2cm}
Approximation of reality based on a set of observations.\\
}
\visible<5>{
\vspace{2cm}
X = observed precipitation of today,\\
\qquad current temperature / humidity / \dots,\\
\qquad physical model output\\
\vspace{0.15cm}
Y = daily precipitation sums\\
}

\vspace{-3.5cm}
\visible<7->{
\textbf{Application/Motivation:}
\begin{itemize}
\item Prediction.
\item Caption of dependencies and interactions.
\end{itemize}
}
\vspace{0.5cm}
\visible<8->{
\textbf{Two main approaches / ``cultures'':}
\begin{itemize}
\item Data modeling.
\item Algorithmic modeling.
\end{itemize}
}
\end{frame}

\begin{frame}[fragile]
\frametitle{Data modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {f(X; par)};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\only<1-6>{
\begin{itemize}
\item Approximate data generating process up to a few unknown processes.
\item Mathematical equations, model assumptions.
\item Data: estimation of parameters.
\end{itemize}
}
\only<7-10>{
\vspace{1cm}
Summary:
\begin{itemize}
\item<8-10> Stochastic model to approximate data generating process.\\
\only<9-10>{$\Rightarrow$ fixed $f$\\}
\item<10> Disadvantage: Risk of incorrect assumptions.
\end{itemize}
}
\only<1>{\vspace{1cm}}

@@

\visible<2-6>{
\begin{minipage}{0.65\textwidth}
\textbf{Example:} 
\only<2-3>{Linear model}\only<4>{Generalized additive model}\only<5-6>{Generalized additive model\\ for location, scale, and shape}
\\
\only<2-3>{
$$\visible<3>{\mathbb{E}(Y|X) = } f(X; \alpha) = \alpha_0 + \alpha_1 \cdot X$$  
\vspace{0.15cm}
}
\only<4>{
$$\mathbb{E}(Y|X) = f(X; \alpha) = \alpha + s(X)$$  
\vspace{0.15cm}
}
\only<5-6>{
\vspace{-0.8cm}
\begin{align*}
\visible<6>{\mathbb{E}(Y|X) = }\mu(X) = \alpha_{\mu} + s_{\mu}(X)\\  
\visible<6>{\sqrt{Var(Y|X)} = }\sigma(X) = \alpha_{\sigma} + s_{\sigma}(X)
\end{align*}
}
\end{minipage}
%\vspace{-0.6cm}
%\hfill
\begin{minipage}{0.3\textwidth}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<2-3>{
<<plot_motivation_GLM_single, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<4>{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\only<5-6>{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
}
\end{minipage}
}
\end{frame}

@@

sp_rain <- partysplit(2L, index = 1:2)
sp_humidity <- partysplit(4L, breaks = 70)

pn <- partynode(1L, split = sp_rain, kids = list(
  partynode(2L, split = sp_humidity, kids = list(
      partynode(3L, info = "4 mm/24h"),
      partynode(4L, info = "9 mm/24h"))),
  partynode(5L, info = "15 mm/24h")))

p <- party(pn, weather)
@
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
nobs <- 120
x <- c(1:nobs)/nobs
ytrue <- ytree <- ytree2 <- ytree3 <- ytree4 <- ytree_up <- yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<0.55) 4 else 4+5*plogis(kappa*(3*(x[i]-0.36)-1))
y <- ytrue + rnorm(nobs,0,2)
y_up <- rep.int(15, length(x)) + rnorm(nobs,0,1)
for(i in 1:nobs) ytree[i] <- if(x[i]<0.7) 4 else 9
for(i in 1:nobs) ytree2[i] <- if(x[i]<0.6) 4 else 9
for(i in 1:nobs) ytree3[i] <- if(x[i]<0.75) 4 else 9
for(i in 1:nobs) ytree4[i] <- if(x[i]<0.55) 4 else 9
ytree_up <- rep.int(15, length(x))
for(i in 1:nobs) yforest[i] <- if(x[i]<0.55) 4 else {if(x[i]<0.6) 5 else {if(x[i]<0.7) 6.5 else {if(x[i]<0.75) 8 else 9}}}
@
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
#%for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@

\begin{frame}[fragile]
\frametitle{Algorithmic modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\draw[<->] (0.1,0.1) .. controls (0.2, -0.4) and (0.4, -0.5) .. (0.8,-0.5);
\draw[<->] (2.7,-0.5) .. controls (3.1, -0.5) and (3.3, -0.4) .. (3.4,0.1);
\node[font=\scriptsize] at (1.75,-0.5) {algorithm};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\visible<2->{
\begin{itemize}
\item Flexible approximation with no/few assumptions %\\
%$\Rightarrow$ no fixed $f$.
\item Data-driven model fitting.
\end{itemize}
}
\only<2>{\vspace{1cm}}
\visible<3->{
\begin{minipage}{0.6\textwidth}
%\vspace{-0.5cm}
\textbf{Example:} \only<3-6>{Tree model\only<3>{\\}\only<4-6>{s}}\only<7->{Forest model}\\
%\vspace{-0.8cm}
\begin{center}
\only<3>{
\vspace{-1cm}
\resizebox{0.6\textwidth}{!}{
<<plot_raintree, fig=TRUE, echo=FALSE, width=4, height=2.7>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.4,0))
paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(p,
     tp_args = list(FUN = identity, width = 9, fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1)
@
}

\vspace{-0.7cm}
\begin{flushleft}
\noindent {\small X = rain now, humdity\\
Y = rain 24h}
\end{flushleft}}
\only<4->{
\vspace{-0.5cm}
\resizebox{0.55\textwidth}{!}{
\begin{tikzpicture}
%\draw[ellipse, draw=black, align=center, scale = 0.7, minimum width=170pt, minimum height = 35pt] (set) at (1, 3) {};
\draw[gray] (1,3.5) ellipse (3 and 0.8);
\begin{scope}
\clip (1,3.5) ellipse (3 and 0.8);
\pgfmathsetseed{7}
\foreach \p in {1,...,200} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
\end{scope}
\visible<5->{
\node[ellipse, draw=black!50!green, align=center, minimum width=80pt, minimum height = 40pt, line width = 3pt] (subset1) at (0.2, 3.5) {};
\visible<7->{
\node[ellipse, draw = blue, align=center, minimum width=100pt, minimum height = 30pt, line width = 3pt] (subset2) at (1.2, 3.6) {};
\visible<9->{
\node[ellipse, draw = orange, align=center, minimum width=120pt, minimum height = 20pt, line width = 3pt] (subset3) at (1.6, 3.2) {};
\visible<6->{
  \node[ellipse, fill=black!50!green, align=center] (n00) at (-1, 2) {};
  \node[ellipse, fill=black!50!green, align=center] (n01) at (-1.25, 1.25) {};
  \draw[-, line width=1pt] (n00) -- (n01);
  \node[rectangle, fill=black!50!green, align=center] (n02) at (-1.5, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n02);
  \node[rectangle, fill=black!50!green, align=center] (n03) at (-1, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n03);
  \node[rectangle, fill=black!50!green, align=center] (n04) at (-0.5, 0.5) {};
  \draw[-, line width=1pt] (n00) -- (n04);
\visible<8->{
  \node[ellipse, fill=blue, align=center] (n10) at (1, 2) {};
  \node[ellipse, fill=blue, align=center] (n11) at (0.75, 1.25) {};
  \draw[-, line width=1pt] (n10) -- (n11);
  \node[rectangle, fill=blue, align=center] (n12) at (0.5, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n12);
  \node[rectangle, fill=blue, align=center] (n13) at (1, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n13);
  \node[rectangle, fill=blue, align=center] (n14) at (1.5, 0.5) {};
  \draw[-, line width=1pt] (n10) -- (n14);
\visible<10->{
  \node[ellipse, fill=orange, align=center] (n20) at (3, 2) {};
  \node[ellipse, fill=orange, align=center] (n21) at (2.75, 1.25) {};
  \draw[-, line width=1pt] (n20) -- (n21);
  \node[rectangle, fill=orange, align=center] (n22) at (2.5, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n22);
  \node[rectangle, fill=orange, align=center] (n23) at (3, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n23);
  \node[rectangle, fill=orange, align=center] (n24) at (3.5, 0.5) {};
  \draw[-, line width=1pt] (n20) -- (n24);
\visible<11->{
\draw[-, line width=1pt] (-1, -0.2) -- (0.2,-0.7);
\draw[-, line width=1pt] (1, -0.2) -- (1,-0.7);
\draw[-, line width=1pt] (3, -0.2) -- (1.8,-0.7);
\node[rectangle, align=center] (combine) at (1, -1) {\large average};
\end{tikzpicture}
\end{center}
\end{minipage}
\hspace{-1cm}
\begin{minipage}{0.4\textwidth}
\begin{center}
\only<3-5>{
%\vspace{-0.29cm}
<<plot_motivation_regtree, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=7, lty = 2)
mtext("rain", side = 4, las = TRUE, at = 15, line = 0.5, cex = 3, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 8, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.4, cex = 3)
%\hspace{-1cm}
\only<6-7>{
<<plot_motivation_randforest1, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=7, lty = 2)
lines(x = x, y = ytree2, col = pal["tree"], lwd=7)
mtext("rain", side = 4, las = TRUE, at = 15, line = 0.5, cex = 3, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 8, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.4, cex = 3)
\only<8-9>{
<<plot_motivation_randforest2, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=7, lty = 2)
lines(x = x, y = ytree2, col = pal["tree"], lwd=7)
lines(x = x, y = ytree3, col = "blue", lwd=7)
mtext("rain", side = 4, las = TRUE, at = 15, line = 0.5, cex = 3, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 8, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.4, cex = 3)
\only<10>{
<<plot_motivation_randforest3, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=7, lty = 2)
lines(x = x, y = ytree2, col = pal["tree"], lwd=7)
lines(x = x, y = ytree3, col = "blue", lwd=7)
lines(x = x, y = ytree4, col = "orange", lwd=7)
mtext("rain", side = 4, las = TRUE, at = 15, line = 0.5, cex = 3, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 8, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.4, cex = 3)
\only<11->{
<<plot_motivation_randforest4, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=7, lty = 2)
lines(x = x, y = yforest, col = pal["forest"], lwd=7)
mtext("rain", side = 4, las = TRUE, at = 15, line = 0.5, cex = 3, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 8, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.4, cex = 3)
\end{center}
\end{minipage}
